---
title: "2024 Big Data Bowl"
author: "Alex Laycock"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(lubridate)
library(janitor)
library(mgcv)
library(fitdistrplus)
library(tweedie)
library(statmod)
library(itsadug)
library(keras)
library(tensorflow)
library(gt)
library(gtExtras)
library(nflfastR)
library(ggimage)

#install_tensorflow(envname = "r-tensorflow")
#install_keras()
# use_condaenv("keras-tf", required = TRUE)
```

``` {r}
# Loading in the data

games <- read.csv("games.csv")
players <- read.csv("players.csv") %>% clean_names() %>%
  dplyr::mutate(ft = str_split_i(height,"-",1) %>% as.numeric(),
                inch = str_split_i(height,"-",2) %>% as.numeric(),
                height_in = ft*12+inch
                )
plays <- read.csv("plays.csv") 
tackles <- read.csv("tackles.csv") 

# Loading in the first week of tracking data to play around with
track1 <- read.csv("tracking_week_1.csv") %>% clean_names()

teams <- nflfastR::teams_colors_logos

# Use dir instead of o to reference player motion
```

What event words are the key words that should be used when determining different factors

``` {r}
track1 %>% pull(event) %>% unique()
# first_contact , tackle, qb_slide , touchdown, qb_sack , fumble - end 
# run, handoff, lateral, snap_direct, pass_arrived, pass_outcome_caught - start
start_triggers <- c("run", "handoff", "lateral", "snap_direct", "pass_arrived", "pass_outcome_caught")
end_triggers <- c("first_contact" , "tackle", "qb_slide" , "touchdown", "qb_sack" , "fumble","out_of_bounds")
```

Keep only tackle and first contact for validation set. 


### Yards Saved Metric and Why It Matters

The metric we've chosen to create for this project is a yards saved metric. As it currently stands, yards saved can only be calculated on plays where there is a missed tackle. These are the plays to investigate to see how prevalent and how big of an impact yards saved has on the game of football. 

``` {r}

num_plays_w_tackle <- tackles %>% dplyr::group_by(gameId,playId) %>%
  summarise(missed_tackle_binomial = ifelse(sum(pff_missedTackle, na.rm = TRUE)> 0,1,0)) %>%
  ungroup() %>% nrow()

missed_tackles <- tackles %>% filter(pff_missedTackle == 1)

num_plays_w_missed_tackle <- missed_tackles %>% distinct(gameId,playId) %>% nrow()

num_plays_w_missed_tackle / num_plays_w_tackle

```

Missed tackles only occurred on 15% of the plays with a tackle involved. 

``` {r}
# track1 %>% inner_join(missed_tackles, by = c("game_id" = "gameId","play_id" = "playId", "nfl_id" = "nflId")) %>% head(20)


missed_tackle_yards <- function(week) {
  # read in
  trackfile <- paste0("tracking_week_",week,".csv")
  track1 <- read_csv(trackfile) %>% clean_names()
  
  # only plays with missed tackles
  track_mt <- track1 %>% inner_join(missed_tackles, by = c("game_id" = "gameId","play_id" = "playId"))
  # only the ball carrier for those plays
  allcar <- track_mt %>%
        inner_join(plays %>% dplyr::select(gameId,playId,ballCarrierId), 
                   by = c("game_id" = "gameId","play_id" = "playId","nfl_id" = "ballCarrierId")) %>%
        setNames(paste0("car_",names(.))) 
  
  # start indicated at first contact
  # end indicated with the end tags: tackle, touchdown, oob
  
  mtsum <- allcar %>%
    dplyr::group_by(car_game_id,car_play_id) %>%
    dplyr::summarise(start_frame = pick(car_event,car_frame_id) %>% filter(car_event == "first_contact") %>% 
                      pull(car_frame_id) %>% min(),
                    end_frame = pick(car_event,car_frame_id) %>% 
                      filter(car_event %in% c("tackle","qb_slide","touchdown",
                                          "qb_sack","fumble","out_of_bounds")) %>%
                      pull(car_frame_id) %>% min(),
                    start_x = pick(car_event,car_x) %>% filter(car_event == "first_contact") %>% 
                      pull(car_x) %>% min(),
                    end_x = pick(car_event,car_x) %>% 
                      filter(car_event %in% c("tackle","qb_slide","touchdown",
                                          "qb_sack","fumble","out_of_bounds")) %>%
                      pull(car_x) %>% min(),
                    diff = abs(end_x - start_x)
                  )
  
  
  return(mtsum)

}

mt_summary <- 1:9 %>% map_dfr(missed_tackle_yards)


# Taking the difference between first contact and tackle as the yards gained lost for a missed tackle
```

Now we have every missed tackle and the difference in yards between the tackle occuring and the first contact. We can analyze the distribution to understand the affect of this better.

``` {r}
mt_summary %>% filter(diff == Inf) %>% nrow() ### errant tagged plays

track1 %>% inner_join(mt_summary %>% filter(diff == Inf), by = c("game_id" = "car_game_id","play_id" = "car_play_id")) %>% View()
# maybe investigate why there is an infinite difference
# those plays don't have a first contact variable, can't assign value to missed yards

mt_summary <- mt_summary %>% filter(diff < Inf)

mt_summary %>%
  ggplot(aes(x = diff)) + geom_density()

mt_summary %>% pull(diff) %>% summary() # Distribution of the yards gained after missed tackles

total_missed_tackle_yds <- mt_summary %>% pull(diff) %>% sum() # Total yards from missed tackles 

total_yards <- plays %>% inner_join(tackles) %>% distinct(gameId,playId, .keep_all = TRUE) %>%
  pull(playResult) %>% sum()

# Percentage of all yards coming from missed tackles
total_missed_tackle_yds / total_yards

```

On plays with a missed tackle, average of 7.7 yards and median of 5.5 yards were gained after the missed tackle. Total of 13,104 yards occurred after missed tackles. Also, for plays where a tackle was made, 15.5%  of the total yards came after a missed tackle. This is a significant amount of yards that resulted from missed tackles. Definitely something that should be accounted for. 

Which teams gave up the most missed tackles and most yards from missed tackles?

``` {r}
mt_plot <- mt_summary %>% 
  left_join(plays, by = c("car_game_id" = "gameId","car_play_id" = "playId")) %>%
  dplyr::group_by(defensiveTeam) %>%
  dplyr::summarise(yds_from_missed_tackles = sum(diff, na.rm = TRUE)) %>%
  ungroup() %>%
  inner_join(teams %>% dplyr::select(team_abbr,team_logo_wikipedia,team_color),
             by = c("defensiveTeam" = "team_abbr")) %>%
  ggplot() +
  geom_bar(aes(x = reorder(defensiveTeam,desc(yds_from_missed_tackles)), y = yds_from_missed_tackles,
               fill = team_color),
           stat = "identity") +
  geom_image(aes(x = reorder(defensiveTeam,desc(yds_from_missed_tackles)), y = -25, 
                 image = team_logo_wikipedia),size = 0.04) +
  scale_fill_identity() +
  xlab("") + ylab("Yards Allowed From Missed Tackles") +
  ggtitle("2022 Week 1-9 Defensive Yards Allowed From Missed Tackles") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"))

ggsave(filename = "mt_plot.png",plot = mt_plot,dpi = 320,
       height = 6, width = 8,
       bg = "white")
# bring in nflverse plotting info to add logos and colors to this plot

```

### Getting the average duration between catch/handoff and tackle

``` {r}
# Create a loop that takes a week of tracking data
# groups all the plays
# returns a data frame with the play id and the time between start and end triggers
# use the ball for only one set of tracking data per play
tr_files <- list.files()[grepl("tracking",list.files())]

carry_duration <- function(row_n) {
  
  
    bbbb <- read.csv(tr_files[row_n]) %>%
      clean_names() %>%
      filter(display_name == "football") %>%
      dplyr::mutate(starts  = ifelse(event %in% start_triggers, "start",NA),
                ends = ifelse(event %in% end_triggers, "end",NA),
                cap_frames = ifelse(starts == "start" | ends == "end", frame_id,NA)) %>%
      dplyr::group_by(game_id,play_id) %>%
      dplyr::summarise(duration = max(cap_frames, na.rm = TRUE) - min(cap_frames, na.rm = TRUE)) %>% 
      ungroup()
  return(bbbb)
}

play_carry_duration <- 1:length(tr_files) %>% map_dfr(carry_duration)

ggplot(play_carry_duration, aes(x = duration / 10)) + geom_density()

play_carry_duration %>% pull(duration) %>% summary()
```

Average time with the ball is 2.7 seconds, to model the tackles with polynomials we will build the validation set with carries of at least 2 seconds (~25th percentile). This will assure that there is adequate data to model the trajectory of the distance between players. 

Filtering the dataset of plays to meet all the conditions for testing

``` {r}
long_runs <- play_carry_duration %>% filter(duration > 20)

# note 1809 plays result in out of bounds

play_testing_set <- inner_join(plays,long_runs, by = c("gameId" = "game_id","playId" = "play_id")) %>%
  filter(preSnapHomeTeamWinProbability > 0.1, preSnapHomeTeamWinProbability < 0.99,
         !grepl("sack",playDescription,ignore.case = TRUE),
         !grepl("touchdown",playDescription,ignore.case = TRUE),
         !grepl("ob ",playDescription,ignore.case = TRUE))
```



### Setting up the data for each carrier - defender pair for modeling

Set up LSTM with multiple batches based on timer duration. Max of 104 hz (timestamps) need to change each pair into length 104. Padding concept, but instead of padding with zeros, pad on the front end with the initial distance possibly. Or use 0 at the end to pad and accomidate for the zeros. Then set the batch length to be 104 to feed into the model each pair. "Masking feature" in keras to ignore 0 value, could use this.


Dont need carrier tackler and non tackler, I want every carrier / defender pair with only the columns I need to run lstm. Changing function to just do that and return below


``` {r}


car_def_pairs_each_week <- function(week) {
  stt <- Sys.time()
  trackfile <- paste0("tracking_week_",week,".csv")
  track1 <- read_csv(trackfile) %>% clean_names()
  
  tackle_involved_id <- tackles %>% pull(nflId)
  
  alldef <- track1 %>% 
    inner_join(plays %>% dplyr::select(gameId,playId,defensiveTeam), 
                   by = c("game_id" = "gameId","play_id" = "playId")) %>%
    filter(display_name != "football",
           club == defensiveTeam
           ) 

  allcar <- track1 %>%
        inner_join(plays %>% dplyr::select(gameId,playId,ballCarrierId), 
                   by = c("game_id" = "gameId","play_id" = "playId","nfl_id" = "ballCarrierId")) %>%
        setNames(paste0("car_",names(.))) 
  

  pair <- inner_join(allcar,alldef, 
                                by = c("car_game_id" = "game_id",
                                       "car_play_id" = "play_id",
                                       "car_frame_id" = "frame_id")) %>%
    dplyr::group_by(car_game_id,car_play_id,nfl_id) %>%
    dplyr::mutate(start = pick(event,car_frame_id) %>% filter(event %in% start_triggers) %>% 
                    pull(car_frame_id) %>% min(),
                  alt_start = pick(event,car_frame_id) %>% filter(event  == "ball_snap") %>% 
                    pull(car_frame_id) %>% min(),
                  start = ifelse(start == Inf,alt_start,start),
                  end = pick(event,car_frame_id) %>% filter(event == "tackle") %>% # changing this from end trigger to tackle in hopes for improvement
                    pull(car_frame_id) %>% min(),
                  dist_to_car = sqrt(((x-car_x)^2) +  ((y-car_y)^2)),
                  timer = car_frame_id - start,
                  unq_play_id = paste0(car_game_id,"-",car_play_id,"-",car_nfl_id,"-",nfl_id)
                  ) %>%
    ungroup() %>%
    filter(car_frame_id >= start,car_frame_id <= end) %>%
    dplyr::select(unq_play_id,timer,dist_to_car)

    ste <- Sys.time()
  print(ste-stt)
  return(pair)

}

carrier_def <- 1:9 %>% map_dfr(car_def_pairs_each_week) # trying for one week just to see how it works

```

I can use keras LSTM to predict one value ahead and iteratively feed the value into a new dataset to predict, for however many values i need. Better than having the model predict many points in advance.

### Initial setup with one LSTM layer

```{r}


set.seed(222)

maxlen <- carrier_def$timer %>% max()

unq <- carrier_def$pair_id %>% unique()

dist_mn <- mean(carrier_def$dist_to_car)
dist_sd <-sd(carrier_def$dist_to_car)



calculate_lags <- function(df, var, lags){
  map_lag <- lags %>% map(~partial(lag, n = .x))
  return(df %>% mutate(across(.cols = {{var}}, .fns = map_lag, .names = "{.col}_lag{lags}")))
}


# how many lags should be used for lstm 
lagz <- 20


padded <- expand_grid(
  unq_play_id = carrier_def$unq_play_id %>% unique(),
  timer = 0:maxlen) %>%
  left_join(carrier_def %>% distinct() %>% dplyr::mutate(dist_to_car = (dist_to_car - dist_mn) / dist_sd)) %>%
  dplyr::select(unq_play_id,timer,dist_to_car) %>%
  replace_na(list(dist_to_car = -dist_mn/dist_sd)) %>%
  dplyr::group_by(unq_play_id) %>%
  calculate_lags(dist_to_car,1:lagz) %>%
  replace(is.na(.),0)

split <- group_initial_split(padded,
                             group = unq_play_id,
                             p = 0.75)

tr_lstm <- training(split) 
te_lstm <- testing(split)


# paddedlist <- split(as.matrix(padded[,-c(1:3)]), 1:nrow(padded))
tr_paddedx <- as.matrix(tr_lstm[,-c(1:3)])
tr_paddedy <- tr_lstm$dist_to_car

tr_paddedx_arr <- array(
    data = as.numeric(unlist(tr_paddedx)),
    dim = c(
        nrow(tr_paddedx),
        lagz,
        1
    )
)

tr_paddedy_arr <- array(
    data = as.numeric(unlist(tr_paddedy)),
    dim = c(
        length(tr_paddedy),
        1,
        1
    )
)

MLmetrics::RMSE(tr_paddedy, lag(tr_paddedy,1,default = 0)) # 0.129 is the baseline RMSE
MLmetrics::RMSE(te_paddedy, lag(te_paddedy,1,default = 0)) # 0.129 is the baseline RMSE

# may want to add in layer_embeddin masking zero

rm(padded,tr_lstm,te_lstm,split)
gc()

model <- keras_model_sequential() %>%
  layer_masking(mask_value = -dist_mn/dist_sd,
                batch_input_shape = c(maxlen+1, lagz, 1) # batch size, timesteps, features
                             ) %>%
  layer_lstm(units = 32,
             #return_sequences = TRUE,
             stateful = TRUE
             ) %>%
  layer_dense(units = 1)

model %>%
  compile(loss = "mse", optimizer = "adam")

summary(model)

################ This is running the model currently 6 hrs per epoch ############

with(tf$device("CPU"),
    rez <- model %>%
      fit(x = tr_paddedx_arr, y = tr_paddedy_arr,
          batch_size = maxlen+1,
          epochs = 10,
          validation_split = 0.2,
          shuffle = FALSE)
)

plot(rez)

################################################################################

te_paddedx <- as.matrix(te_lstm[,-c(1:3)])
te_paddedy <- te_lstm$dist_to_car

te_paddedx_arr <- array(
    data = as.numeric(unlist(te_paddedx)),
    dim = c(
        nrow(te_paddedx),
        lagz,
        1
    )
)

te_paddedy_arr <- array(
    data = as.numeric(unlist(te_paddedy)),
    dim = c(
        length(te_paddedy),
        1,
        1
    )
)
with(tf$device("CPU"),
model_forecast <- model %>%
  predict(te_paddedx_arr, batch_size = maxlen+1)
)

MLmetrics::RMSE(model_forecast,te_paddedy)



with(tf$device("CPU"),
     model %>% evaluate(te_paddedx_arr,te_paddedy_arr, batch_size = maxlen+1) # 1.37 with one layer
)
```


### Adding complexity with second layer & saving model weights and only going to 10 epochs based on training

Need to save the fitted weights of the model to a callback file in order to load these in and make predictions on the testing sets going forward. Retraining the model since it was not saved before. Using the loss plot, no real improvements after 10 epochs so we will stop at 10.


``` {r}

model <- keras_model_sequential() %>%
  layer_masking(mask_value = -dist_mn/dist_sd,
                batch_input_shape = c(maxlen+1, lagz, 1) # batch size, timesteps, features
                             ) %>%
  layer_lstm(units = 32,
             return_sequences = TRUE,
             dropout = 0.1,
             recurrent_dropout = 0.5,
             stateful = TRUE
             ) %>%
  layer_lstm(units = 32,
             #return_sequences = TRUE,
             stateful = TRUE
             ) %>%
  layer_dense(units = 1)

model %>%
  compile(loss = "mse", optimizer = "adam")

summary(model)

checkpoint_path <- "lstm_2l_training/cp.ckpt"
checkpoint_dir <- fs::path_dir(checkpoint_path)

# Create a callback that saves the model's weights
cp_callback <- callback_model_checkpoint(
  filepath = checkpoint_path,
  save_weights_only = TRUE,
  verbose = 1
)


with(tf$device("CPU"),
    rez2 <- model %>%
      fit(x = tr_paddedx_arr, y = tr_paddedy_arr,
          batch_size = maxlen+1,
          epochs = 10,
          validation_split = 0.2,
          shuffle = FALSE,
          callbacks = list(cp_callback)
          )
)


load_model_weights_tf(model,checkpoint_path) # with proper trained weights

with(tf$device("CPU"),
  model %>% evaluate(te_paddedx_arr,te_paddedy_arr, batch_size = maxlen + 1) ### 0.0624 loss
)


```

Very good model results with 0.06 MSE on a held out test set.

Need to make predictions for entire dataset. Then develop loops for subsequent test sets following each for the next 100 frames. Then we can run all of the next tackler stats.

``` {r}

last_ts_lagged <- carrier_def %>% distinct() %>% 
  dplyr::mutate(dist_to_car = (dist_to_car - dist_mn) / dist_sd) %>%
  dplyr::select(unq_play_id,timer,dist_to_car) %>%
  replace_na(list(dist_to_car = -dist_mn/dist_sd)) %>%
  dplyr::group_by(unq_play_id) %>%
  calculate_lags(dist_to_car,1:lagz) %>%
  replace(is.na(.),0) %>%
  dplyr::group_by(unq_play_id) %>%
  filter(timer == max(timer)) %>%
  ungroup()

last_ts_lagged %>%
  ggplot(aes(x = dist_to_car)) +
  geom_density()


```


Build a function to make predictions on each of the last time frames and predict 100 timestamps in advance. Naturally will predict out the number of steps from the batch size, 144. 144 frames is too long, need to predict ahead the next 42 frames. This covers 95% of cases.

``` {r}

set_random_seed(444)
load_model_weights_tf(model,checkpoint_path) # with proper trained weights

build_100_frame_predictions <- function(rowx) {
  
  lf_num <- demo_set$timer[rowx]
  
  id <- demo_set$unq_play_id[rowx]

  matt <- as.matrix(demo_set[rowx,-c(1:3)])
  
  arr <- array(
      data = as.numeric(unlist(matt)),
      dim = c(
          1,
          lagz,
          1
      )
  )
  
  
  
  with(tf$device("CPU"),
  next_pred <- (model %>%
    predict(arr, verbose = 0))[1,1]
  )
  
  preds <- numeric()
  preds[1] <- next_pred
  
  cur_frame <- c(next_pred,as.numeric(unlist(matt))[1:19])
  for (i in 2:42) {
    

    newarr <- array(
      data = cur_frame,
        dim = c(
            1,
            lagz,
            1
            )
    )

   with(tf$device("CPU"),
    new_pred <- (model %>% predict(newarr, verbose = 0))[1,1]
        )
    
   preds[i] <- new_pred
   
   old_vals <- cur_frame[1:19]
   cur_frame <- c(new_pred,old_vals)
    
    
  }
  
  full_preds <- data.frame(unq_play_id = id,
             timer = seq(lf_num+1,lf_num+42),
             dist_to_car = as.numeric(preds)
  )
  
  
  return(full_preds)


  
}

# subsetting the last_ts dataframe for only the last week

wk9gms <- read.csv("tracking_week_9.csv") %>% 
  clean_names() %>%
  pull(game_id) %>%
  unique()

demo_set <- last_ts_lagged %>%
  dplyr::mutate(game_id = unq_play_id %>% str_split_i("-",1)) %>%
  filter(game_id %in% wk9gms) %>%
  dplyr::select(-game_id)

demo_predictions <- 1:nrow(demo_set) %>% map(build_100_frame_predictions, .progress = TRUE)

demo_predictions_fin <- bind_rows(demo_predictions)
```

Predictions and example of the model will be used on only the last week of the season, not enough size/memory to make predictions for all the weeks. These will be used going forward to determine yards saved and exploratory data tables. 

Need to take out the actual tacklers from the predicted demo set. Want to keep only the nontacklers to look at the next hypothetical tackle situation. With the predictions, now for each play we will look at the next possible collision with a defender, distance less than 1.6 yards away. For each play we want to return one row. The row will show the time stamp of the earliest predicted tackle, or return that no tackle was predicted.

``` {r}

tack_gr_each_play <- tackles %>%
  filter(tackle == 1 | assist == 1 | forcedFumble == 1) %>%
  dplyr::group_by(gameId,playId) %>%
  dplyr::summarise(tack_all_ids = list(nflId)) %>%
  ungroup()


scaled_cut <- (1.6 - dist_mn) / dist_sd

frame_sum <- demo_predictions_fin %>%
  dplyr::mutate(game_id = str_split_i(unq_play_id,"-",1),
                play_id = str_split_i(unq_play_id,"-",2),
                car_nfl_id = str_split_i(unq_play_id,"-",3),
                def_nfl_id = str_split_i(unq_play_id,"-",4)) %>%
  inner_join(tack_gr_each_play %>% dplyr::mutate(across(everything(),as.character)),
               by = c("game_id" = "gameId","play_id" = "playId")) %>%
  dplyr::group_by(game_id,play_id) %>%
  filter(!(def_nfl_id %in% tack_all_ids)) %>%
  ungroup() %>%
  dplyr::group_by(game_id,play_id,timer) %>%
  summarise(pred_tackle = ifelse(min(dist_to_car) < scaled_cut,1,0),
            pred_closest_dist = min(dist_to_car),
            pred_tackler_id = pick(dist_to_car,def_nfl_id) %>% filter(dist_to_car == pred_closest_dist) %>%
              pull(def_nfl_id)) %>%
  ungroup()
View(frame_sum)


play_sum <- frame_sum %>%
  dplyr::group_by(game_id,play_id) %>%
  summarise(new_tackle_occurs = ifelse(sum(pred_tackle) > 0,1,0),
            new_tackle_frame = case_when(
              new_tackle_occurs == 1 ~ pick(timer,pred_tackle) %>% filter(pred_tackle == 1) %>% 
                pull(timer) %>% min(),
              TRUE ~ max(timer)),
            orig_tackle_frame = min(timer)-1,
            diff_in_frames = new_tackle_frame - orig_tackle_frame,
            new_tackler_id = pick(new_tackle_frame,pred_tackler_id) %>% filter(timer == new_tackle_frame) %>%
              pull(pred_tackler_id)
              )

play_sum %>%
  ggplot(aes(x = diff_in_frames)) +
  geom_density()
  
play_sum %>% pull(diff_in_frames) %>% summary()

```





# no assisted tackles, no missed tackles, just solo clean tackle
``` {r, echo = FALSE}
clean_tackles <- tackles %>%
  filter(tackle == 1, assist == 0, pff_missedTackle == 0)

carrier_only_data <- function(week) {
  
  trackfile <- paste0("tracking_week_",week,".csv")
  track1 <- read_csv(trackfile) %>% clean_names()
  

  
  allcar <- track1 %>%
        inner_join(plays %>% dplyr::select(gameId,playId,ballCarrierId), 
                   by = c("game_id" = "gameId","play_id" = "playId","nfl_id" = "ballCarrierId")) %>%
        inner_join(players %>% dplyr::select(nfl_id, height_in,weight)) %>%
        setNames(paste0("car_",names(.))) %>%
        # inner_join(play_testing_set %>% dplyr::select(gameId,playId),
        #        by = c("car_game_id" = "gameId", "car_play_id" = "playId")) %>%
        dplyr::group_by(car_game_id,car_play_id) %>%
    dplyr::mutate(start = pick(car_event,car_frame_id) %>% filter(car_event %in% start_triggers) %>% 
                    pull(car_frame_id) %>% min(),
                  end = pick(car_event,car_frame_id) %>% filter(car_event == "tackle") %>% # changing this from end trigger to tackle in hopes for improvement
                    pull(car_frame_id) %>% min(),
                  timer = car_frame_id - start
                  ) %>%
    ungroup() %>%
    filter(car_frame_id >= start, car_frame_id <= end) %>%
    dplyr::select(car_game_id,car_play_id,timer,car_x)

  return(allcar)

}

carrier <- 9 %>% map_dfr(carrier_only_data)
```




### Projecting hypothetical yards

``` {r}
# building a gam model and returning prediction based on data
carr_traj <- function(subset_df) {
  # subset df should consist of car_x, timer and predicted frames
  mod <- gam(car_x ~ s(timer, bs = "cr", k = 3), data = subset_df)
  summary(mod)
  pred_x <- predict.gam(mod, newdata = data.frame(timer = subset_df$new_tackle_frame[1]))
  return(pred_x)
}


# putting together carrier and tackler with next hypothetical tackler
rb_other_traj <- carrier %>%
  dplyr::mutate(across(car_game_id:car_play_id,as.character)) %>%
  inner_join(play_sum, by = c("car_game_id" = "game_id","car_play_id" = "play_id")) %>%
  dplyr::group_by(car_game_id,car_play_id) %>%
  dplyr::mutate(x_from_traj = carr_traj(pick(car_x,timer,new_tackle_frame)) ) %>%
  ungroup()

yds_saved_db <- rb_other_traj %>%
  dplyr::group_by(car_game_id,car_play_id) %>%
  filter(timer == max(timer)) %>% # last row for each play
  ungroup() %>%
  dplyr::mutate(yds_saved = abs(x_from_traj - car_x)) %>%
  inner_join(tackles %>%
              filter(tackle == 1) %>%
              dplyr::group_by(gameId,playId) %>%
              slice(1) %>%
              ungroup() %>%
              mutate(across(gameId:playId,as.character)) %>% 
              setNames(paste0("orig_",names(.))) , 
        by = c("car_game_id" = "orig_gameId", "car_play_id" = "orig_playId")) %>%
  dplyr::select(car_game_id,car_play_id,orig_nflId,orig_tackle_frame,
                new_tackler_id,new_tackle_frame,diff_in_frames,yds_saved)

yds_saved_db %>%
  ggplot(aes(x = yds_saved)) +
  geom_density()

yds_saved_db %>%
  ggplot(aes(x = diff_in_frames,y = yds_saved)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```

We now have yards saved for all the demo plays and can see the clear relationship between the predicted difference in time between the next hypothetical collision and yards saved. 

### Evaluating yards saved by team and by player for the demo set

``` {r}
#player
pl_tab_viz <- yds_saved_db %>%
  inner_join(plays %>%
               mutate(across(gameId:playId,as.character)) %>% 
               dplyr::select(gameId,playId,quarter:defensiveTeam), 
             by = c("car_game_id" = "gameId","car_play_id" = "playId")) %>%
  inner_join(players %>% 
               dplyr::select(nfl_id,position,display_name) %>%
               setNames(paste0("orig_tackler_",names(.))), by = c("orig_nflId" = "orig_tackler_nfl_id")) %>%
  inner_join(teams %>% dplyr::select(team_abbr,team_logo_wikipedia), 
             by = c("defensiveTeam" = "team_abbr")) %>% 
  dplyr::group_by(car_game_id,team_logo_wikipedia,defensiveTeam,possessionTeam,orig_nflId,
                  orig_tackler_display_name,orig_tackler_position) %>%
  summarise(tackles = n(),
            total_yds_saved = round(sum(yds_saved),2),
            yds_saved_per_tackle = round(total_yds_saved / tackles,2)) %>%
  ungroup() %>%
  arrange(car_game_id,defensiveTeam,desc(total_yds_saved)) %>%
  gt() %>%
  tab_header(title = "2022 Week 9 Yards Saved Individual Leaders") %>%
  cols_label(
    car_game_id ~ "Game ID",
    team_logo_wikipedia ~ "Logo",
    defensiveTeam ~ "Team",
    possessionTeam ~ "Opponent",
    orig_nflId ~ "ID",
    orig_tackler_display_name ~ "Tackler",
    orig_tackler_position ~ "POS",
    tackles ~ "Tackles",
    total_yds_saved ~ "Total Yards Saved",
    yds_saved_per_tackle ~ "Yards Saved Per Tackle"
  ) %>%
  gt_img_rows(columns = team_logo_wikipedia) %>%
  data_color(
    columns = c(tackles,total_yds_saved,yds_saved_per_tackle),
    method = "numeric",
    palette = "PiYG",
    na_color = "white"
  ) %>%
  opt_interactive(
    use_search = TRUE,
    use_filters = TRUE,
    use_resizers = TRUE,
    use_highlight = TRUE,
    use_compact_mode = TRUE,
    use_text_wrapping = FALSE,
    use_page_size_select = TRUE
  ) %>%
  opt_align_table_header(align = "center") %>%
  cols_width(
    car_game_id ~ px(150),
    defensiveTeam ~ px(80),
    possessionTeam ~ px(80),
    orig_nflId ~ px(80),
    orig_tackler_display_name ~ px(150),
    orig_tackler_position ~ px(80),
    team_logo_wikipedia ~ px(80),
  ) %>%
  cols_align(align = "center")

  gtsave(pl_tab_viz, filename = "pl_tab_viz.html")

```


``` {r}
#team
  
tm_tab_viz <- yds_saved_db %>%
  inner_join(plays %>%
               mutate(across(gameId:playId,as.character)) %>% 
               dplyr::select(gameId,playId,quarter:defensiveTeam), 
             by = c("car_game_id" = "gameId","car_play_id" = "playId")) %>%
  inner_join(teams %>% dplyr::select(team_abbr,team_logo_wikipedia), 
             by = c("defensiveTeam" = "team_abbr")) %>% 
  dplyr::group_by(car_game_id,team_logo_wikipedia,defensiveTeam) %>%
  summarise(tackles = n(),
            total_yds_saved = round(sum(yds_saved),2),
            yds_saved_per_tackle = round(total_yds_saved / tackles,2)) %>%
  ungroup() %>%
  arrange(car_game_id,defensiveTeam,desc(total_yds_saved)) %>%
  gt() %>%
  tab_header(title = "2022 Week 9 Yards Saved By Team") %>%
  cols_label(
    car_game_id ~ "Game ID",
    defensiveTeam ~ "Team",
    team_logo_wikipedia ~ "Logo",
    tackles ~ "Tackles",
    total_yds_saved ~ "Total Yards Saved",
    yds_saved_per_tackle ~ "Yards Saved Per Tackle"
  ) %>%
  gt_img_rows(columns = team_logo_wikipedia) %>%
  data_color(
    columns = c(tackles,total_yds_saved,yds_saved_per_tackle),
    method = "numeric",
    palette = "PiYG",
    na_color = "white"
  ) %>%
  opt_interactive(
    use_search = TRUE,
    use_filters = TRUE,
    use_resizers = TRUE,
    use_highlight = TRUE,
    use_compact_mode = FALSE,
    use_text_wrapping = FALSE,
    use_page_size_select = TRUE,
    page_size_default = 32
  ) %>%
  opt_align_table_header(align = "center") %>%
  cols_width(
    car_game_id ~ px(150),
    defensiveTeam ~ px(80),
    team_logo_wikipedia ~ px(80)
  ) %>% 
  cols_align(align = "center")

  gtsave(tm_tab_viz, filename = "tm_tab_viz.html")
```

From the individual summarized resuts you can see the importance of tackles added to the DBs and LBs where the margins for making a tackle are much higher (more to lose with missed tackles). This further highlights the inefficiency of using tackles itself as a metric, considering the contrast in value assigned across the position groups. 


### Yards saved by defensive position group

``` {r}

general_pos_table <- tibble(
  orig_tackler_position = c("CB","DB","DE","DT","FS","ILB","MLB","NT","OLB","SS"),
  pos_group = c("DB","DB","DL","DL","DB","LB","LB","DL","LB","DB")
)

general_pos_viz <- yds_saved_db %>%
  inner_join(plays %>%
               mutate(across(gameId:playId,as.character)) %>% 
               dplyr::select(gameId,playId,quarter:defensiveTeam), 
             by = c("car_game_id" = "gameId","car_play_id" = "playId")) %>%
  inner_join(players %>% 
               dplyr::select(nfl_id,position,display_name) %>%
               setNames(paste0("orig_tackler_",names(.))), by = c("orig_nflId" = "orig_tackler_nfl_id")) %>%
  inner_join(general_pos_table) %>%
  dplyr::group_by(pos_group) %>%
  summarise(tackles = n(),
            total_yds_saved = round(sum(yds_saved),2),
            yds_saved_per_tackle = round(total_yds_saved / tackles,2)) %>%
  ungroup() %>%
  pivot_longer(cols = tackles:yds_saved_per_tackle) %>%
  ggplot(aes(x = factor(pos_group, levels = c("DB","LB","DL")), y = value, fill = name)) +
  geom_bar(position = "dodge",stat = "identity") +
  facet_wrap(~name, scales = "free") +
  theme_bw() + 
  ggtitle("Tackle Metrics Position Group Breakdown",
          subtitle = "2022 Week 9") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title = element_blank(),
        legend.title = element_blank(),
        strip.background = element_rect(alpha("lightgrey",0.5)))

ggsave(filename = "general_pos_viz.png",
       plot = general_pos_viz,
       dpi = "retina",
       scale = 2)
  
```

Grouping the positions into the three levels of the defense (DL,LB, and DB), the importance of tackles at each level can be clearly distinguished. As you would expect, the DB group is the most critical in saving yards on tackles. When watching games, you can see missed tackles from DB's resulting in huge gains for the offense. Most of the time, DL's are making tackles with others, or there are many other defenders around the ball carrier. In these situations, the DL making the tackle would not be saving many yards, given that there are other defenders who can swiftly step in to make the tackle if needed.



Could add percentage of tackles that occured after the 